\assignment

### 2017-03-18 project/paper.rnw
Described the UXO site.

<<packages, message = FALSE, cache = FALSE>>=
library(spatstat)

set.seed(83534)
@

@@

Previously this semester, we worked with an example of a Poisson process on
the unit square \(\{(x,y):0<x<1,0<y<1\}\) with the log-linear intensity
surface
\begin{equation*}
\log(\lambda(x,y)) = 5x + 2y.
\end{equation*}
I will continue this example, fitting the model
\begin{equation*}
\log(\lambda(x,y)) = \beta_{0} + \beta_{1}x + \beta_{2}y
\end{equation*}
first using the process observed over the full region (the entire unit
square), and then using the process observed in a subregion treating
\(\{(x,y):0.6<x<0.9,0.6<y<0.9\}\) as unobserved. This unobserved section is a
region of relatively high intensity, so if the estimation or prediction
methods are sensitive to the observation window I expect omitting this
section to have a large effect.

I will use both estimated models to predict on
\(\{(x,y):-0.5<x<1.5,-0.5<y<1.5\}\). Figure~\ref{fig:hw4_plot} shows the
realization of this process that I will use for both model fits.

@@

@@

After estimating the models, I use \texttt{spatstat}'s \texttt{predict.ppm}
method to predict the intensity surface on a \(128 \times 128\) lattice of
locations in the region
\(\{(x,y):-0.5<x<1.5,-0.5<y<1.5\}\)~(Figure~\ref{fig:hw4_fitplot}). As
expected from the similarity of the coefficient estimates, the predicted
log-intensity surfaces for both models show very similar linear trends~(top
row of the figure). At a glance the prediction standard errors for both
models appear similar to the predicted intensities, as seen in the middle
row of the figure~(shown on a logarithmic scale for visibility). In a linear
model setting we would expect the prediction standard error to increase with
the distance from the observations, but no such trend is apparent here; even
in the bottom left of the prediction region the standard errors decrease as
the predicted intensity decreases. However, it is illuminating to plot the
predictor's relative standard error (the ratio of the standard error to the
predicted intensity, bottom row of the figure). This ratio is not constant,
and in fact it is lowest where the highest intensity of events was observed.

@@

easy site, realization 2,000, 390 ft spacing

>>>>

<<packages, message = FALSE, cache = FALSE>>=
library(parallel)
library(spatstat)

set.seed(c(83534, 251, -18691, 283, -2526, 782), kind = 'L\'Ecuyer-CMRG')
@

@@

@@

Previously this semester, we worked with an example of a Poisson process on
the unit square \(\{(x,y):\Sexpr{min(simulate_win$xrange)} < x
< \Sexpr{max(simulate_win$xrange)}, \Sexpr{min(simulate_win$yrange)} < y
< \Sexpr{max(simulate_win$yrange)}\}\) with the log-linear intensity surface
\begin{equation*}
\log(\lambda(x,y)) = 5x + 2y.
\end{equation*}
I will continue this example, fitting the model
\begin{equation*}
\log(\lambda(x,y)) = \beta_{0} + \beta_{1}x + \beta_{2}y
\end{equation*}
first using the process observed over the full region (the entire unit
square), and then using the process observed in a subregion treating
\(\{(x,y): \Sexpr{min(omit_win$xrange)} < x < \Sexpr{max(omit_win$xrange)},
\Sexpr{min(omit_win$yrange)} < y < \Sexpr{max(omit_win$yrange)}\}\) as
unobserved. This unobserved section is a region of relatively high intensity,
so if the estimation or prediction methods are sensitive to the observation
window I expect omitting this section to have a large effect.

I will use both estimated models to predict on
\(\{(x,y): \Sexpr{min(predict_win$xrange)} < x
< \Sexpr{max(predict_win$xrange)}, \Sexpr{min(predict_win$yrange)} < y
< \Sexpr{max(predict_win$yrange)}\}\). Figure~\ref{fig:hw4_plot} shows the
realization of this process that I will use for both model fits.

@@

After estimating the models, I use \texttt{spatstat}'s \texttt{predict.ppm}
method to predict the intensity surface on a \(128 \times 128\) lattice of
locations in the region \(\{(x,y): \Sexpr{min(predict_win$xrange)} < x
< \Sexpr{max(predict_win$xrange)}, \Sexpr{min(predict_win$yrange)} < y
< \Sexpr{max(predict_win$yrange)}\}\)~(Figure~\ref{fig:hw4_fitplot}). As
expected from the similarity of the coefficient estimates, the predicted
log-intensity surfaces for both models show very similar linear trends~(top
row of the figure). At a glance the prediction standard errors for both
models appear similar to the predicted intensities, as seen in the middle
row of the figure~(shown on a logarithmic scale for visibility). In a linear
model setting we would expect the prediction standard error to increase with
the distance from the observations, but no such trend is apparent here; even
in the bottom left of the prediction region the standard errors decrease as
the predicted intensity decreases. However, it is illuminating to plot the
predictor's relative standard error (the ratio of the standard error to the
predicted intensity, bottom row of the figure). This ratio is not constant,
and in fact it is lowest where the highest intensity of events was observed.

@@

<<uxoexampledata>>=
# Path to my writing project root directory.
vspuxo <- file.path('..', '..', 'vspuxo')

# Load my functions for working with UXO data.
source(file.path(vspuxo, 'rfns', 'data_functions.r'))
source(file.path(vspuxo, 'rfns', 'sampling_functions.r'))
source(file.path(vspuxo, 'rfns', 'spatial_functions.r'))

# Define the window for the UXO site.
sitewindow <- owin(poly = cbind(x = c(1564294, 1564495, 1556870, 1557126),
                                y = c(535421, 541130, 541085, 535576)))
@

In my writing project, I evaluated ordinary kriging of moving average
intensity values on a simulated tank training range. I will use the same site
for this project. The site is a
\Sexpr{sprintf('%.1f', area(sitewindow) / 43560)} acre quadrilateral region
with homogeneous Poisson process producing background anomalies with an
intensity of 100 anomalies per acre, and two regions of concentrated munitions
use generated as Poisson processes with bivariate Gaussian intensity. The
total intesnsity at the center of each concentrated munitions use region is
300 anomalies per acre~(Figure~\ref{fig:truth}).

<<truth, cache = TRUE, fig.width = 3.75, fig.pos = 't', fig.cap = 'True intensity at the simulated UXO site.'>>=
x <- seq(1556870, 1564495, by = 20)
y <- seq(535421, 541130, by = 20)
intense.mat <- matrix(100, nrow = length(x), length(y))
for(i in seq_along(x)){
  for(j in seq_along(y)){
    intense.mat[i, j] <- 100 + gauss.elliptic(x[i]+10, y[j]+10,
        mu.x = 1558400, mu.y = 540000,
        s.a = 800/(2*qnorm(0.995)), s.b = 1200/(2*qnorm(0.995)),
        r = pi/6, maxrate = 200
      ) + gauss.elliptic(x[i]+10, y[j]+10,
        mu.x = 1562000, mu.y = 537000,
        s.a = 2000/(2*qnorm(0.995)), s.b = 900/(2*qnorm(0.995)),
        r = 0, maxrate = 200
      )
  }
}
intense.im <- im(t(intense.mat), x+10, y+10, unitname = c('foot', 'feet'))

par(mar = c(0, 0, 1, 3))
plot(intense.im, main = 'True Intensity Surface', zlim = c(100, 300), las = 2)
plot(setminus.owin(owin(c(1556870, 1564495), c(535421, 541130)),
                   sitewindow), col = 'white', add = TRUE)
mtext('Anomalies per Acre', 4, line = 2)
@

easy site, realization 2,000, 390 ft spacing, 6 ft width
